\textbf{SPECFEM3D}\\ \textbf{User Manual}

### 2011-01-27 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates movie_volume info in SPECFEM3D manual

When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
flag turned on, the code outputs \texttt{moviedata??????} files in
the \texttt{in\_out\_files/OUTPUT\_FILES} directory. The files are in a fairly complicated
binary format, but there is a program provided to convert the
output into more user friendly formats:

\begin{description}
\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
Then type `\texttt{make} \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT}'
and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
will prompt the user for input parameters.

@@

>>>>

When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
flag turned on, the code outputs \texttt{moviedata??????} files in
the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
in the main parameter file \texttt{Par\_file} that control the output of these moviedata files 
(see section \ref{cha:Main-Parameter} for details):
\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity, 
\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
In order to output additionally shakemaps, you would set
the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.

The files are in a fairly complicated
binary format, but there is a program provided to convert the
output into more user friendly formats:
\begin{description}
\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
Then type `\texttt{make} \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT}'
and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
will prompt the user for input parameters.

@@

\subsection{\label{sub:Movie-Volume}Movie Volume}

When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
%that control the region in space and time that is saved. These are:
%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
The output is saved by each processor
at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
For all domains, the velocity field is output to files: 
\texttt{\small proc??????\_velocity\_X\_it??????.bin},
\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
and $\nabla \times \bf{v}$, get stored as well: 
\texttt{\small proc??????\_div\_it??????.bin},
\texttt{\small proc??????\_curl\_X\_t??????.bin},
\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
The files denoted 
\texttt{\small proc??????\_div\_}~\\
\texttt{\small glob\_it??????.bin} and
\texttt{\small proc??????\_curl\_glob\_it??????.bin} 
are stored on the global points only, all the other arrays
are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to 
the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
\par\end{centering}

\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
snapshots of vertical velocity components at different times.}

\label{fig:movie.volume}
\end{figure}

To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
to combine the data from all slices into one mesh file. To compile it in the root directory, type:
\begin{lyxcode}
{\footnotesize make~combine\_vol\_data~}{\footnotesize \par}
\end{lyxcode}
which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
usage of this executable, type `./xcombine\_vol\_data` without arguments.
As an example, you would use
\begin{lyxcode}
{\footnotesize cd~bin/~}{\footnotesize \par}
{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
\end{lyxcode}
to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
We next convert the \texttt{.mesh} file into the VTU (Unstructured
grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
located in directory \texttt{utils/Visualization/Paraview/}, for example:
\begin{lyxcode}
{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
\end{lyxcode}
Notice that this Perl script uses a program \texttt{mesh2vtu} in the
\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
run-time library for its execution. Therefore, make sure you have
them properly set in the script according to your system.
